Supplementary Materials

### 2020-06-02 analysis/analysis.Rmd
Modelling different types of symmetry.

```{r, fig.cap = "Symmetries present in each group"}
sym_dat <- read_csv("../subgroups/symmeries_in_group.csv") %>%
  mutate(group = str_to_upper(group))

>>>>

```{r, fig.cap = "Symmetries present in each group."}
sym_dat <- read_csv("../subgroups/symmeries_in_group.csv") %>%
  mutate(group = str_to_upper(group))

### 2020-06-02 analysis/analysis2.Rmd
Modelling different types of symmetry.

I have set `echo=FALSE` so that most of the code chunks will not display. Please refer to the .Rmd file for source code.

@@

### Fit Model to Psychophysical Data

```{r threshold_fit_model, cache=TRUE, echo=FALSE}
summary(m_threshold)
```

### EEG Control Data

We will also fit models to the control data. As we can see from Figure 2.4, the group differences are much smaller.

```{r fig.cap="The density plot shows the distribution of the empirical data, while the blue line shows the 66% and 95% prediction intervals.", cache=TRUE}

m_eeg_par_odd <- brm(data = d_eeg_par_odd, 
                     rms ~ wg-1 + (1|subject),
                     family = "lognormal", 
                     prior = prior, iter = n_iter,
                     control = list(adapt_delta = 0.99))

my_samples$par_odd <- get_model_samples(m_eeg_par_odd)
my_models$par_odd <- get_model_predictions(my_samples$par_odd)
plt1 <- plot_model_output(my_models$par_odd, d_eeg_par_odd, 
                  "odd-harmonic data from \nparietal electrodes", "orangered3", "threshold (ms)")

m_eeg_occ_even <- brm(data = d_eeg_occ_even, 
                      rms ~ wg-1 + (1|subject),
                      family = "lognormal", 
                      prior = prior, iter = n_iter,
                      control = list(adapt_delta = 0.99))

# Extract samples from model
my_samples$occ_even <- get_model_samples(m_eeg_occ_even)
my_models$occ_even <- get_model_predictions(my_samples$occ_even)
plt2 <- plot_model_output(my_models$occ_even, d_eeg_occ_even, 
                  "even-harmonic data from \noccipital electrodes", "orangered3", "threshold (ms)")

plt1 + plt2 + plot_annotation(title = 'Model Fits for the ')
```

# Subgroup Comparisons

We will now compute the difference between sub- and super-groups.

```{r subgroup_comparisons}
source("scripts/get_subgroup_comparisons.R")

subgroup_comp$eeg <- get_subgroup_comparisons2(my_samples$eeg)
subgroup_comp$threshold <- get_subgroup_comparisons2(my_samples$threshold,)
subgroup_comp$occ_even <- get_subgroup_comparisons2(my_samples$occ_even)
subgroup_comp$par_odd  <- get_subgroup_comparisons2(my_samples$par_odd)
```

## Primary EEG Data

Finally, we calculate the probability that the RMS difference between subgroup and supergroup is larger than zero given the data. This information is then binned so we can colour in the posterior density plots.

```{r define_comparison_summary_function}
comparison_summary <- function(subgroup_comp_dat, greater, measurement_lab) {
  # compute probabilties
  subgroup_comp_dat %>%
    group_by(key, index, normal) %>%
    summarise(
      mean_value = mean(value),
      prob_one_tailed = if_else(greater == TRUE, mean(value > 0), mean(value < 0)))  %>%
    mutate(
      prob_one_tailed = if_else(prob_one_tailed==0, 0.01, prob_one_tailed),
      measurement = measurement_lab,
      index_labs = as.numeric(as.factor(index)),
      lab_cols = v_cols[index_labs],
      prob_cat = cut(prob_one_tailed, c(0, 0.5, 0.75, 0.9, 0.99, 1))) -> comp_summary_dat

  return(comp_summary_dat)
}
```

Now we will use it!

```{r, fig.height=8, fig.cap="Distributions of the difference in mean log(rms) between sub- and super-groups. The index of each relationship is indicated by the colour of the y-axis label. The fill of the density plots indicated the probability of the difference being greater than zero."}
# compute probabilties
comp_summary$eeg <- comparison_summary(subgroup_comp$eeg, TRUE, 'rms')

subgroup_comp$eeg <- full_join(
  subgroup_comp$eeg, 
  comp_summary$eeg, 
  by = c("key", "index", "normal"))

# Finally, plot everything:
p1 <- plot_comparisons(
  filter(subgroup_comp$eeg, index == 2),
  filter(comp_summary$eeg, index == 2), 
  1)

p2 <-plot_comparisons(
  filter(subgroup_comp$eeg, index != 2),
  filter(comp_summary$eeg, index != 2), 
  2)

both_plts <- cowplot::plot_grid(p1, p2, ncol = 2, align="hv")
both_plts
```

## Psychophysical Data

We can do the same for the display duration thresholds from our psychophysics experiment. Here we are looking for the opposite effect, namely that display larger are larger for subgroups than for supergroups (see main paper), so we calculate the probability that differences in duration are smaller than zero.

```{r ,fig.height=8, fig.cap="Distributions of the difference in mean log display duration threshold between sub- and super-groups. The index of each relationship is indicated by the colour of the y-axis label. The fill of the density plots indicated the probability of the difference being less than zero."}

# compute probabilties
comp_summary$threshold <- comparison_summary(subgroup_comp$threshold, FALSE, 'threshold')

subgroup_comp$threshold <- full_join(
  subgroup_comp$threshold, 
  comp_summary$threshold, 
  by = c("key", "index", "normal"))

# Finally, plot everything:
p1 <- plot_comparisons(
  filter(subgroup_comp$threshold, index == 2),
  filter(comp_summary$threshold, index == 2), 
  1, TRUE)

p2 <- plot_comparisons(
  filter(subgroup_comp$threshold, index != 2),
  filter(comp_summary$threshold, index != 2), 
  2, TRUE)

both_plts <- cowplot::plot_grid(p1, p2, ncol = 2, align="hv")
both_plts
```

## Control EEG Data

We will now do exactly the same with the control data (odd harmonic data from parietal electrodes and even harmonic data from occipital electrodes)

```{r fig.height=8, echo = FALSE, fig.cap="Distributions of the difference in mean log(rms) between sub- and super-groups. The index of each relationship is indicated by the colour of the y-axis label. The fill of the density plots indicated the probability of the difference being greater than zero."}

# compute probabilties
comp_summary$occ_even <- comparison_summary(subgroup_comp$occ_even, TRUE, 'rms')

subgroup_comp$occ_even <- full_join(
  subgroup_comp$occ_even, 
  comp_summary$occ_even, 
  by = c("key", "index", "normal"))

p1 <- plot_comparisons(
  filter(subgroup_comp$occ_even, index == 2),
  filter(comp_summary$occ_even, index == 2), 
  1)
p2 <- plot_comparisons(
  filter(subgroup_comp$occ_even, index != 2),
  filter(comp_summary$occ_even, index != 2), 
  2)

both_plts <- cowplot::plot_grid(p1, p2, ncol = 2, align="hv") 
both_plts + plot_annotation(title = 'Odd harmonic data from parietal electrodes')

# compute probabilties
comp_summary$par_odd <- comparison_summary(subgroup_comp$par_odd, TRUE, 'rms')

subgroup_comp$par_odd <- full_join(
  subgroup_comp$par_odd, 
  comp_summary$par_odd, 
  by = c("key", "index", "normal"))

p1 <- plot_comparisons(
  filter(subgroup_comp$par_odd, index == 2),
  filter(comp_summary$par_odd, index == 2), 
  1)
p2 <- plot_comparisons(
  filter(subgroup_comp$par_odd, index != 2),
  filter(comp_summary$par_odd, index != 2), 
  2)

both_plts <- cowplot::plot_grid(p1, p2, ncol = 2, align="hv")
both_plts + plot_annotation(title = 'Even harmonic data from occipital electrodes')
```

## Summary

We can summarise the subgroup comparison plots above by plotting ROC curves for each of our four measurements (Figure 3.5).

```{r, fig.cap="This figure shows how many of our 64 comparisons are classed as having a greater-than-zero difference (less-than-zero for the display durations) for difference thresholds. between 0.5 an 1.0."}

count_comparisons <- function(comparisons, p) {
  x <- sum(comparisons$prob_one_tailed >= p)
  return(x)
}

df <- tibble()

for (p in seq(0.5, 0.99, 0.01)) {

  df <- bind_rows(df, map_dbl(comp_summary,  count_comparisons, p))
}

df$p <- seq(0.5, 0.99, 0.01)

filter(df, p == 0.95)

df %>% gather("measurement", "number", -p) %>%
  mutate(proportion = number/nrow(subgroups)) %>%
  ggplot(aes(x = p, y = proportion, colour = measurement)) + geom_point() + geom_path() +
  theme_minimal() + ggthemes::scale_color_ptol(labels = c("primary EEG", "sensor control EEG", "harmonic control EEG", "psychophysical threshold")) +
  scale_y_continuous(limits = c(0, 1)) +
  xlab("probability of difference given data") + ylab("proportion of subgroup comparisons")
```

If we take p=0.95 as our cut-off, we can see that the subgroup relations are preserved in 56/64 = 87.5% and 49/64 = 76.6% of the comparisons for the primary EEG and display durations repectively. This compares to the $32/64= 50.0% and 22/64 = 34.4% for the control EEG conditions.

# Additional Analysis

## Replicatation of Kohler et al (2016)

We can lok at the groups that only contain orientation, and see if we obtain the parametric response as documented in Kohler et al. (2016). We one might predict based on those results, we get a corresponding pattern of display duration thresholds, with the time taken to detect the symmetry dreasing as we increase the amount of rotational symmetry.

Curiosuly, we do not have a large difference betwee P4 and P6 in the eeg data.

```{r, fig.cap = "The effect of increasing the amount of rotational symmetry.", fig.height=3}
my_samples$eeg %>% select(P2, P3, P4, P6) %>%
  pivot_longer(names_to = "wallpaper group", cols = c(P2, P3, P4, P6)) %>% 
  mutate(exp = "eeg") %>%
  bind_rows(
    my_samples$threshold %>% select(P2, P3, P4, P6) %>% 
      pivot_longer(names_to = "wallpaper group", cols = c(P2, P3, P4, P6)) %>%
      mutate(exp = "display thresholds")) %>% 
  ggplot() + aes(x = value, fill = `wallpaper group`) + 
    geom_density(alpha = 0.35) + 
    facet_wrap(~ exp, scale = "free_y") + 
  theme_minimal() + 
  scale_fill_viridis_d()
```

## Index and Normality

Subgroup relations can be classified by their index, and by whether they are normal or not. Here we investigate the extent to which these two variables can account for the variation between the subgroup relationships.

```{r, cache=TRUE}
my_models <- list()
comp_summary$eeg$index <- as.numeric(as.character(comp_summary$eeg$index))
comp_summary$threshold$index <- as.numeric(as.character(comp_summary$threshold$index))

my_models$eeg <- brm(data = comp_summary$eeg, mean_value ~ index * normal)
my_models$disp <- brm(data = comp_summary$threshold, mean_value ~ index * normal)
summary(my_models$eeg)
summary(my_models$disp)
```

We can see that the index of the subgroup relationship has an effect on both the difference in log(rms) and the difference in log(display duration): relationships with a higher index lead to larger differences.

```{r, fig.cap = "The effect of index and normality on log(rms) and log(ms)."}
comp_summary$eeg %>% 
  modelr::data_grid(index, normal) %>%
  add_fitted_draws(my_models$eeg, n = 100) %>%
  mutate(model = "primary eeg") -> d_plt_eeg

comp_summary$threshold %>% 
  modelr::data_grid(index, normal) %>%
  add_fitted_draws(my_models$disp, n = 100) %>%
  mutate(model = "psychophysical threshold") -> d_plt_disp

d_summary <- bind_rows(comp_summary$threshold %>% mutate(model = "psychophysical threshold"),
                       comp_summary$eeg %>% mutate(model = "primary eeg"))

bind_rows(d_plt_eeg, d_plt_disp) %>%
  ggplot(aes(x =  index, y = .value, color = as.factor(normal), group = interaction(.draw, normal))) +
  geom_line(alpha = .15) +
  facet_wrap(~ model, scales="free_y") + 
  geom_jitter(
    data = d_summary, 
    aes(x =  index, y = mean_value, colour = as.factor(normal), group = 1), width = 0.1, height = 0, alpha = 0.33) +
  theme_minimal() + 
  ggthemes::scale_colour_ptol(labels = c("non-normal", "normal")) + 
  scale_x_continuous("index", breaks = c(2,3,4,6,8,12)) +
  scale_y_continuous("value") +
  theme(panel.grid.minor = element_blank(), legend.title = element_blank()) 
```

## Correlation Between Primary EEG data and Psychophysical Thresholds

Finally, we will investigate whether there is a correlation between the our primary EEG measure (rms amplitude of odd harmonics over occipital cortex) and our display duration thresholds. As our two different measures come from different samples of participants, we are unable to do a direct comparison. However, we can use the results of the models discussed in Section 3 and check for a correlation between the predicted values of the two measures.

```{r compare_comparisons, cache=TRUE}
bind_rows(
  select(comp_summary$eeg, measurement, key, mean_value),
  select(comp_summary$threshold, measurement, key, mean_value)) %>%
  spread(measurement, mean_value) -> d

m <- brm(data = d, threshold ~ rms)

bayes_R2(m)
```

We can see that although the correlation is relatively weak, our confidence interval indicates that we can be reasonably positive that $R^2>0$ (i.e, 95\% credible interval is 0.026 - 0.302).

```{r fig.cap="Scatter plot showing the correlation between our two measures. Each line is a sample from the posterior of a Bayesian linear regression."}
d %>% 
  modelr::data_grid(rms = seq(-0.2, 0.8, 0.1)) %>%
  add_fitted_draws(m, n = 100) %>%
  ggplot(aes(x = rms, y = threshold, color = as.factor(index))) +
  geom_line(aes(y = .value, group = .draw), alpha = .1, colour = "thistle4") +
  geom_point(data = d) +
  theme_minimal() + 
  ggthemes::scale_colour_ptol("index")
```

>>>>

This will be merged into the main file later

@@

## Modelling by symmetry

```{r, cache=TRUE}
d_eeg %>% rename(group = "wg") %>% left_join(sym_dat) -> dat2

m_eeg2 <- brm(data = dat2, 
             rms ~ rotation + reflection + glide + (1|subject),
             family = "lognormal", 
             prior = prior, iter = n_iter,
             control = list(adapt_delta = 0.99))

```{r, cache=TRUE}
d_dispthresh %>% rename(group = "wg") %>% left_join(sym_dat) -> dat3

m_disthresh2 <- brm(data = dat3, 
             threshold ~ rotation + reflection + glide + (1|subject),
             family = "lognormal", 
             prior = prior, iter = n_iter,
             control = list(adapt_delta = 0.99))

```{r}
summary(m_eeg2)

summary(m_disthresh2)
```
